---
title: "HW2"
format: html
editor: visual
---

## Task 1:

Making a vector for the 20 pre-treatment BP values

```{r}
pre <- c(120, 151, 125, 126, 115, 132, 132, 129, 134, 139, 127, 122, 127, 135, 133, 128, 147, 138, 140, 132) 
pre
```

Making a vector for the 20 post-treatment BP values

```{r}
post <- c(127, 145, 135, 122, 115, 122, 123, 126, 126, 129, 132, 146, 120, 114, 121, 120, 128, 120, 133, 115)
post
```

Creating list names for subjects

```{r}
names(pre) <- paste("Subject", 1:20, sep = "_") 
pre

names(post) <- paste("Subject", 1:20, sep = "_") 
post
```

Creating vector for the difference between pre and post BP

```{r}
diff_bp <- pre - post
diff_bp
```

Calculating mean difference in BP

```{r}
mean_bp <- mean(diff)
mean_bp
```

Determining which patients had a decrease in BP from pre to post conditions using the which() function

```{r}
which(diff_bp>0)
```

Using a Boolean as an indexing vector to return only variables in diff_bp that had a positive change

```{r}
positive_change <- c(diff_bp>0)
diff_bp[positive_change]
```

Calculating the mean decrease in BP for patients with a positive change in bp

```{r}
PC_mean <- mean(diff_bp[positive_change]) 
PC_mean
```

## Task 2:

Creating a data frame for patient, pre_bp, post_bp, and diff_bp

```{r}
patient <- 1:20
pre_bp <- c(pre)
post_bp <- c(post)


bp_df <- data.frame(patient, pre_bp, post_bp, diff_bp)
bp_df
```

Return bp_df rows where the diff_bp is negative

```{r}
neg_bp_df <- bp_df[bp_df$diff_bp < 0,]
neg_bp_df
```

Create new column in data frame corresponding to TRUE if the post_bp is less than 120 and FALSE if it is not.

```{r}

bp_df$post_120 <- c(post_bp < 120)
bp_df
```

Print of data frame

```{r}
knitr::kable(bp_df)
```

## Task 3:

Creating a data frame for pre-treatment and post-treatment placebo for subjects 1-10 including a new column for TRUE/FALSE

```{r}
pre_placebo <- c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)
post_placebo <- c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)
placebo_df <- data.frame("Patient" = c(1:10),
                        "pre_placebo" =  c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130), 
                        "post_placebo" = c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124))
                        
placebo_df$placebo_120 <- c(post_placebo < 120)
placebo_df$placebo_diff <- c(pre_placebo - post_placebo)
placebo_df
```

Creating list that combines the treatment and placebo data frames

```{r}
bp_list <- list(placebo = placebo_df, treatment = bp_df)
bp_list
```

Checking the list was properly made by looking at the structure and attributes

```{r}
str(bp_list)
attributes(bp_list)
```

Accessing element one in the list with 3 different types of syntax

```{r}
bp_list[1]
bp_list[[1]]
bp_list$placebo

```

Using bp_list to access the pre_bp column in the placebo data frame

```{r}
bp_list[["placebo"]]$pre_placebo
```
